---
title: "Day 1 - Take home exercise"
format: 
  html: 
    navbar: false
    df-print: paged
    code-copy: true
---

## Dataset

Data `covid_cases.rds` under `R-cafe/day1/data/covid_cases.rds` is used for this exercise.

```{=html}
<details>
  <summary>
    Data dictionary
  </summary>
```
| Column                                             | Description                                              |
|-----------------------------------|-------------------------------------|
| `date`                                             | date of the case report                                  |
| 211 columns with the format `cases_{country_code}` | number of reported cases of each country and archipelago |

Country codes of interest (i.e. Codes for countries needed for the exercise)

| Code  | Country name             |
|-------|--------------------------|
| `chn` | China                    |
| `usa` | United States of America |
| `vnm` | Vietnam                  |
| `sgp` | Singapore                |

</details>

## Work space

Before working on this exercise, you should have completed the following checklist

::: callout-note
### Checklist

On Github Desktop

-   Cloned [R-cafe repository](https://github.com/OUCRU-Modelling/R-cafe)

-   Created your own branch on the repository

-   Checkout your own branch

On RStudio

-   Make an R project from cloned R-cafe directory

-   Have renv setup
:::

::: callout-note
## Answer script

You are expected to save your answer under `R-cafe/day1/takehome.R`
:::

## Exercises

### Task 1: Data import

Read data from `R-cafe/day1/data/covid_cases.rds`

### Task 2: Simple computations using dataset

Compute the following

-   The **earliest** and **latest** date of data report in the dataset and assign them to 2 variables `first_report_date`, `last_report_date`

-   Create new column `case_global` for `covid_cases`, which represents the total cases across every country **per report day**.

-   Create new column `percent_chn` in `covid_cases`, which represents the percentage of global cases that China's cases account for per report day**.**

::: {.callout-tip collapse="true"}
### Tip: compute sum of each row

Use function `rowSums()` to compute sum of case columns row wise

**Example**

```{r}
example_table <- data.frame(
  a = c(1:3), b = c(4:6)
)
# show example_table
example_table

# see output of rowSums
rowSums(example_table)
```
:::

::: {.callout-tip collapse="true"}
### Tip: create new column/ update column values

To create/update a column, use the syntax `data[["colname"]] <- values_for_col` where length of `values_for_col` must be equal to number of rows

**Example**

```{r}
example_table <- data.frame(
  a = c(1:3), b = c(4:6)
)

# create new column new_col
example_table[["new_col"]] <- c(7,9,11)
example_table
```
:::

### Task 3: Create a function

Complete the following:

-   Create a function `compute_percent` that

    -   takes the dataset and the [**country code**]{.underline} (i.e. the 3 characters after `cases_` in column name) as input.

    -   return the percentage of the global cases that the given country's cases account for per report day.

-   Use that function to create 3 new columns for `covid_cases`: `percent_vnm`, `percent_use`, `percent_sgp`

-   Print the final `covid_cases` selecting only the following columns `date`, `percent_chn`, `percent_vnm`, `percent_use`, `percent_sgp`

::: callout-tip
### Tip: joining strings

Use function `paste()` to join multiple strings together

**Example**

```{r}
# joining 2 strings
paste("cases_", "chn", sep = "")

# joining 3 strings
paste("hello", "world,", "welcome to R-cafe", sep = " ")
```
:::

### Task 4: Generate data summary

Complete the following:

-   Use the function `skim()` from package `skimr` to generate summary for the following countries: China, Vietnam, USA, Singapore

-   Describe notable observations from each country from the result of `skim()`

### Task 5: Peer review

Complete the following:

-   Push your solution to Github with the commit message `"day 1 take home solution your_name"`

-   Create an issue commenting on at least 1 other participant's solution
